---
title: tidyeval, some musings on dplyr::filter
author: Sebastian Sauer
date: '2022-02-09'
slug: tidyeval-some-musings-on-dplyr-filter
categories:
  - rstats
tags:
  - tutorial
---



```{r global-knitr-options, include=FALSE}
  knitr::opts_chunk$set(
  fig.pos = 'H',
  fig.asp = 0.618,
  fig.align='center',
  fig.width = 5,
  out.width = "100%",
  fig.cap = "", 
  dpi = 300,
  # tidy = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  fig.show = "hold")
```




# Programming with the tidyverse


... Is not exactly self-evident. It actually requires some head wrapping, at
least in my experience. 
In this post, we are exploring some aspects on programming when filtering rows.
Let's see.



# Setup

```{r}
library(tidyverse)
```




# Some filtering chunk


Let's say we would like to filter observations according to some variable and a given threshold in some data set:



```{r}
mtcars %>% 
  filter(hp > 200)
```


Of course, that's a toy problem, but for the sake of the tidyverse programming thing,
it may be a helpful start to keep things straight.

Now, say, instead of `200`, we abstract it to `threshold`, a variable giving the filter value:


```{r}
threshold <- 200

mtcars %>% 
  filter(hp > threshold)
```

Works! So `filter` is able to digest a variable, at least at the right hand side of the filter expression.


# Now let's abstract it to a function


```{r}
filter_fun <- function(data, threshold){
  
  data %>% 
    filter(hp > threshold)
}
```


Let's try it:


```{r}
filter_fun(mtcars, 200)
```


Works!



# Actract variable to be filtered



```{r}
filter_fun <- function(data, col, threshold){
  
  data %>% 
    filter({{col}} > threshold)
}
```



Let's try it:


```{r}
filter_fun(mtcars, hp, 200)
```

Works!


We can make use of the "curly curly" operator as in `{{col}}` to explain to the tidyverse function
that what now comes is a variable, and *not* a column name in the present data frame.





