---
title: Overlaying facetted histograms with normal curve using ggplot2
author: Sebastian Sauer
date: '2021-06-23'
slug: overlaying-facetted-histograms-with-normal-curve-using-ggplot2
categories:
  - rstats
tags:
  - visualization
  - dataviz
---



# Overlaying histograms with a normal curve


Overlaying a histogram (possibly facetted) is not something far fetched when analyzing data. Surprisingly, it appears (to the best of my knowledge) that there's no comfortable out-of-the-box solution in ggplot2, although it can be of course achieved with some lines of code.  Here's my take.


# Setup


```{r}
library(tidyverse)
```


# Some data

```{r}
d <- read_csv("https://vincentarelbundock.github.io/Rdatasets/csv/openintro/speed_gender_height.csv")
```


```{r}
glimpse(d)
```

```{r}
d %>% 
  slice_head(n = 5)
```



# Preparing data

We'll use a "total" histogram for the whole sample, to that end, we'll need to remove the grouping information from the data.

```{r}
d2 <-
  d |> 
  select(-gender)
```


Here's a data set with summary data:

```{r}
d_summary <-
  d %>% 
  group_by(gender) %>% 
  summarise(height_m = mean(height, na.rm = T),
            height_sd = sd(height, na.rm = T))

d_summary
```


# Plot it

```{r}
d |> 
  ggplot() +
  aes() +
  geom_histogram(aes(y = ..density.., x = height, fill = gender)) +
  facet_wrap(~ gender) +
  geom_histogram(data = d2, aes(y = ..density.., x = height), 
                 alpha = .5) +
  stat_function(data = d_summary %>% filter(gender == "female"),
                fun = dnorm,
                #color = "red",
                args = list(mean = filter(d_summary, 
                                          gender == "female")$height_m,
                            sd = filter(d_summary, 
                                        gender == "female")$height_sd)) +
  stat_function(data = d_summary %>% filter(gender == "male"),
                fun = dnorm,
                #color = "red",
                args = list(mean = filter(d_summary, 
                                          gender == "male")$height_m,
                            sd = filter(d_summary, 
                                        gender == "male")$height_sd)) +
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  labs(title = "Facetted histograms with overlaid normal curves",
       caption = "The grey histograms shows the whole distribution (over) both groups, i.e. females and men") +
  scale_fill_brewer(type = "qual", palette = "Set1")
   
```