---
title: Using quizzes on markdown html sites
author: Sebastian Sauer
date: '2024-01-18'
slug: using-quizzes-on-markdown-html-sites
categories:
  - rstats
tags:
  - markdown
  - quiz
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Motivation

As a teacher, I often write exercises for my students and post the exercises on my [Datenwerk site](https://datenwerk.netlify.app/).

As many questions are of multiple choice type it would come handy to have a quiz function.

I first hoped that Quarto markdown would supply such a feature out of the box.
However, this is not (yet) the case.
The good news is that it can be achieved with some simple html/js code.


# Simple quiz implementation using HTML and JS


```{html, eval = FALSE}
<form id="quizForm">
    <h4>Behauptung: Das 50%-PI [.6,.8] ist das plausiblere.</h4>
    <label>
      <input type="radio" name="q1" value="Ja"> Ja
    </label>
    <label>
      <input type="radio" name="q1" value="Nein"> Nein
    </label>
    <label>
      <input type="radio" name="q1" value="Keine Antwort möglich"> Keine Antwort möglich
    </label>

    <button type="button" onclick="submitQuiz()">Antworten</button>
</form>

  <script>
  function submitQuiz() {
      var selectedOption = getSelectedOption("q1");
      var correctAnswer = "Ja";

      // Display feedback
      if (selectedOption === correctAnswer) {
        alert("Richtig!");
      } else {
        alert("Falsch. Die richtige Antwort lautet *Ja*.");
      }
    }

    function getSelectedOption(questionName) {
      var radioButtons = document.getElementsByName(questionName);
      for (var i = 0; i < radioButtons.length; i++) {
        if (radioButtons[i].checked) {
          return radioButtons[i].value;
        }
      }
      return null;
    }
 </script>

```


# Demo

See here some live demos:

- [Example 1](https://datenwerk.netlify.app/posts/verteilungen-quiz-02/verteilungen-quiz-02)
- [Example 2](https://datenwerk.netlify.app/posts/verteilungen-quiz-16/verteilungen-quiz-16)





# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


