---
title: Preparing German weather data
author: Sebastian Sauer
date: '2022-07-24'
slug: preparing-german-weather-data
categories:
  - rstats
tags:
  - data
  - datawrangling
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
library(glue)
```


# Motivation

In this post, we'll prepare official German weather data. All data are accessed from [DWD](https://www.dwd.de/DE/Home/home_node.html).

More specifically, the DWD (Deutscher Wetterdienst, German Weather Service) provides an open data section <https://opendata.dwd.de/>, from where the data were fetched.


# Licence

The licence can be read [here](https://opendata.dwd.de/README.txt), in short, the data are provided free of charge for everybody's use.

At [this subpage](https://www.dwd.de/EN/service/copyright/copyright_node.html) however, it is stated that the materials are copy right protected and that source must be declared like this:

>   Datenbasis: Deutscher Wetterdienst, eigene Elemente ergänzt.

(Data Source: Deutscher Wetterdiesnt, own elements added).

And, furthermore:

>   When speaking of spatial data, this also includes any location-related weather and climate information presented on our open web pages.




# It's a playful approach

I'm not pretending to do full-fetched meterology research.
What's being done is a rather playful appraoch to derive some basic insights noentheless.


# Download data


##  Air temperature means

Here's the URL stem, indicating a folder where the data can be found:

```{r url-stem}
url_stem <- "https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/air_temperature_mean/"
```


There's one file for each month, providing the mean air temperature for the German Bundesländer:

```{r build-file-names}
months <- str_pad(1:12, width = 2, side = "left", pad = "0")
months
month_temp_data <- glue("regional_averages_tm_{months}.txt")
month_temp_data
```
Let's glue the whole path:

```{r build-urls}
air_temp_path <- paste0(url_stem,month_temp_data)
air_temp_path
```


# Download multiple files and bind them together rowwise

First, check:

Here's the first path:

```{r}
air_temp_path[1]
```



```{r readpath1}
d <- read_delim(air_temp_path[1], delim = ";", skip = 1)
head(d)
```


However, for this file, some columns are of a *different* type:




```{r}
d <- read_delim(air_temp_path[11], delim = ";", skip = 1)
head(d)
```

See col `Monat`.

So let's better read all cols as character, and post-process the numnbers out of the strings.




```{r check2}
d <- read_delim(air_temp_path[11], delim = ";", skip = 1, col_types = cols(.default = "c")) %>% select(-`...20`)

head(d)
```


First, download all files and append to one tibble:

```{r download-all-files}
d_air_temp <- 
  air_temp_path %>% 
  map_df(~ read_delim(., skip = 1, col_types = cols(.default = "c"))) %>% 
  select(-`...20`)
```


 
Second, let's parse the numbers out of the strings.

```{r parse-numbers}
d_air_temp2 <-
  d_air_temp %>% 
  mutate(across(.cols = everything(),
                 .fns = parse_number))

d_air_temp2 %>% 
  head() %>% 
  gt::gt()
```

# Format to long 

```{r format-lng}
d_air_temp3 <-
  d_air_temp2 %>% 
  pivot_longer(cols = -c(Jahr,, Monat), names_to = "region", values_to = "temp") %>% 
  rename(year = Jahr, month = Monat)
```


# Save to disk

I'm saving to my personal folder. Your folder path will be different.

```{r}
write_csv(d_air_temp3, file = "/Users/sebastiansaueruser/github-repos/datasets/csv/air_temp_de_dev_DWD.csv")
```


# Reference temperature

In this data set, the temperatures are denotes in devations from a long term average.
I think this reference time is 1961-1990, [building on this source](https://www.dwd.de/DE/leistungen/besondereereignisse/verschiedenes/20210119_neue_referenzperiode.html;jsessionid=B502C793E6F64A251E0902D8F85B9FA8.live31081?nn=16102), but after 2021, the reference period has been flipped to 1991-2020.



# Debrief

This is nice stuff, however, the temperatures provided are *deviations* from a long term average, which may be of interest at times, but not so tangible as actual temperatures.



# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


