---
title: Programmatically plotting with ggplot2
author: Sebastian Sauer
date: '2022-09-28'
slug: programmatically-plotting-with-ggplot2
categories:
  - rstats
tags:
  - visualization
  - dataviz
  - tidyverse
  - ggplot2
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Setup

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
library(easystats)  # comfort in stats
```



```{r}
data(mtcars)
```


In essence,
we want to build this kind of plot programmatically:



```{r}
mtcars %>% 
  ggplot(aes(x=hp)) +
  geom_histogram()
```




# Let's go

## Way 1

Let's use unquoted variable names.

```{r}
plot_descriptives0 <- function(data, var) {
  

  data %>% 
   ggplot(aes(x = {{var}})) +
   geom_histogram()
}

plot_descriptives0(mtcars, hp)
```



## Way 2

Somewhat more elaborated, but same idea:

```{r}
plot_descriptives1 <- function(data, var, group) {
  
  var_string <- deparse(substitute(var))
  
  data %>% 
    select({{var}}, {{group}}) %>%
    drop_na() %>% 
    group_by({{group}}) %>%
    mutate({{group}} := as.factor({{group}})) %>%  
    describe_distribution(iqr = FALSE, range = TRUE, quartiles = TRUE) %>% 
    select(Variable, Mean, SD, .group) %>% 
    mutate(group = str_remove(.group, "^.+=")) %>% 
    select(-.group) %>% 
    #pivot_longer(-c(Variable, group)) %>% 
    ggplot(aes(x = group)) +
    geom_errorbar(aes(ymin = Mean-(2*SD), ymax = Mean+(2*SD)), width = .1) +
    geom_point2(aes(y = Mean), alpha = .7, size = 3) +
    facet_wrap(~ Variable) +
    labs(caption = "Error bars show mean ± 2SD",
         title = paste0("Variable: ", var_string)) +
    coord_flip() +
    theme(legend.position = "bottom")
}
```




```{r}
plot_descriptives1(mtcars, hp, cyl)
```




This line gives us the variable name as a string:
```{r}
deparse(substitute(var))
```

More specifically:

```{r}
deparse(substitute(mtcars$hp))
```





## Way 2



This time, we have variables as *strings*

```{r}
plot_descriptives2 <- function(data, var) {
  data %>% 
    ggplot(aes_string(x = var)) +
    geom_histogram() +
    labs(title = paste0("Variable: ", var))
}
```


However, `aes_string` is deprecated.

```{r}
plot_descriptives2(mtcars, "hp")
```



## Way 3

Here we use `sym`

```{r}
?sym
```


>   sym() and syms() take strings as input and turn them into symbols.

See:

```{r}
sym("hp")
```


!! is used to evaluate a symbol, ie., gives back the "real" column, such as `hp`.




>   The !! operator unquotes its argument. It gets evaluated immediately in the surrounding context.


```{r}
plot_descriptives3 <- function(data, var) {
  
  var_enquo <- sym(var)
  
  data %>% 
    ggplot(aes(x = !!var_enquo)) +
    geom_histogram()
}

plot_descriptives3(mtcars, "hp")

```



## Way 4

We can swqp the "bang-bang" !! in favor of "curly-curly" :

```{r}
plot_descriptives4 <- function(data, var) {
  
  var_enquo <- sym(var)
  
  data %>% 
    ggplot(aes(x = {{var_enquo}})) +
    geom_histogram()
}

plot_descriptives4(mtcars, "hp")
```


# Further reading

There are many sources, here are some:

- https://stackoverflow.com/questions/53381764/what-does-sym-do-regarding-tidyeval
- https://www.rdocumentation.org/packages/modeltime.gluonts/versions/0.1.0/topics/tidyeval


# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


