---
title: Adjust labels in ggplot
author: Sebastian Sauer
date: '2024-02-25'
slug: adjust-labels-in-ggplot
categories:
  - rstats
tags:
  - ggplot2
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
library(see)  # okabeito_colors
```


# Data

```{r}
data("mariokart", package = "openintro")
```


# Unadjusted labels

```{r}
mario_quantile <- 
mariokart %>% 
  filter(total_pr < 100) %>% 
  summarise(q25 = quantile(total_pr, .25),
            q50 = quantile(total_pr, .50),
            q75 = quantile(total_pr, .75))

mario_quantile <- 
  mariokart %>% 
  filter(total_pr < 100) %>% 
  reframe(qs = quantile(total_pr, c(.25, .5, .75)))

mariokart %>% 
  filter(total_pr < 100) %>%  
  ggplot(aes(x = total_pr)) +
  geom_histogram() +
  geom_vline(xintercept = mario_quantile$qs) +
  annotate("label", x =  mario_quantile$qs, y = 0, label =  mario_quantile$qs) +
  annotate("label", x =  mario_quantile$qs, y = Inf, label =  c("Q1", "Median", "Q3")) 
```


# Adjusted labels manually

`vjust = 2` does the trick:

```{r}
mario_quantile <- 
mariokart %>% 
  filter(total_pr < 100) %>% 
  summarise(q25 = quantile(total_pr, .25),
            q50 = quantile(total_pr, .50),
            q75 = quantile(total_pr, .75))

mario_quantile <- 
  mariokart %>% 
  filter(total_pr < 100) %>% 
  reframe(qs = quantile(total_pr, c(.25, .5, .75)))

mariokart %>% 
  filter(total_pr < 100) %>%  
  ggplot(aes(x = total_pr)) +
  geom_histogram() +
  geom_vline(xintercept = mario_quantile$qs) +
  annotate("label", x =  mario_quantile$qs, y = 0, label =  mario_quantile$qs) +
  annotate("label", x =  mario_quantile$qs, y = Inf, label =  c("Q1", "Median", "Q3"), vjust = 2) 
```

# Adjust labels automatically


Using `vjust = "top"` (or "bottom") does the trick in an automatic, comfortable way.


```{r}
mariokart %>% 
  filter(total_pr < 100) %>%  
  ggplot(aes(x = total_pr)) +
  geom_histogram() +
  geom_vline(xintercept = mario_quantile$qs) +
  annotate("label", x =  mario_quantile$qs, y = 0, label =  mario_quantile$qs, vjust = "bottom") +
  annotate("label", x =  mario_quantile$qs, y = Inf, label =  c("Q1", "Median", "Q3"), vjust = "top") 
```


Similarly, one case use `hjust = "left"` for horizontal alignment.

# Expanding the limits

Let's consider a second example, where it helped to expand the limits of the coordinates.

In this example, we use Weather data from the German DWD  [i.e., Deutschen Wetterdienst, DWD](https://www.dwd.de/DE/Home/home_node.html) .

After some [data wrangling, which will not discussed here](https://data-se.netlify.app/2022/07/24/preparing-german-weather-data/), there's a nice data set waiting for further scrutiny.


```{r}
wetter_path <- "https://raw.githubusercontent.com/sebastiansauer/Lehre/main/data/wetter-dwd/precip_temp_DWD.csv"
wetter <- read.csv(wetter_path)
```


Some data preparation:



```{r}
wetter <-
  wetter %>% 
  mutate(after_1950 = year > 1950) %>% 
  filter(region != "Deutschland")  # ohne Daten für Gesamt-Deutschland
```

Here's a linear model for our weather data:

```{r}
lm_wetter_region <- lm(temp ~ region, data = wetter)
```


```{r}
wetter_summ <- 
  wetter %>% 
  group_by(region) %>% 
  summarise(temp = mean(temp)) %>% 
  mutate(id = 0:15) %>% 
  ungroup() %>%
  mutate(grandmean = mean(temp),
         delta = temp - grandmean)

wetter_summ %>%  
  ggplot(aes(y = region, x = temp)) +
  theme_minimal() +
  geom_label(aes(label = paste0("b", id),
                 x = grandmean + delta), 
             vjust = 1,
             size = 2,
             color = okabeito_colors()[1]) +
  geom_vline(xintercept = coef(lm_wetter_region)[1], linetype = "dashed", color = okabeito_colors()[2]) +
  
  #coord_cartesian(xlim = c(7, 10), ylim = c(0, 16))  +
 
  annotate("label",
           y = 1,
           x = coef(lm_wetter_region)[1],
           vjust = 1,
           label = paste0("b0"),
           #size = 6,
           color = okabeito_colors()[2]) +
  annotate("point", 
           y = 1, 
           x = coef(lm_wetter_region)[1], 
           color = okabeito_colors()[2],
           #vjust = 1,
           size = 4) +
  geom_segment(aes(yend = region, xend = temp), 
               x = coef(lm_wetter_region)[1], 
               color = okabeito_colors()[1])  +
    geom_point() +
  labs(y = "",
       x = "Temperatur")
```

Behold the label right at the bottom of the diagram, "b0".
It is not displayed properly.

One way to get it back into the picture is bay extending the limits of the y-axis:


```{r}

wetter_summ %>%  
  ggplot(aes(y = region, x = temp)) +
  theme_minimal() +
  geom_label(aes(label = paste0("b", id),
                 x = grandmean + delta), 
             vjust = 1,
             size = 2,
             color = okabeito_colors()[1]) +
  geom_vline(xintercept = coef(lm_wetter_region)[1], linetype = "dashed", color = okabeito_colors()[2]) +
  
 coord_cartesian(xlim = c(7, 10), ylim = c(0, 16))  +  # back in the game!
 
  annotate("label",
           y = 1,
           x = coef(lm_wetter_region)[1],
           vjust = 1,
           label = paste0("b0"),
           #size = 6,
           color = okabeito_colors()[2]) +
  annotate("point", 
           y = 1, 
           x = coef(lm_wetter_region)[1], 
           color = okabeito_colors()[2],
           #vjust = 1,
           size = 4) +
  geom_segment(aes(yend = region, xend = temp), 
               x = coef(lm_wetter_region)[1], 
               color = okabeito_colors()[1])  +
    geom_point() +
  labs(y = "",
       x = "Temperatur")
```


# Duckdive the problem: tinyfy the label

It may feel like cheating, but, hey, problem solved: just decrease the size of the label.


```{r}

wetter_summ %>%  
  ggplot(aes(y = region, x = temp)) +
  theme_minimal() +
  geom_point() +
  annotate("point", 
           y = 1, 
           x = coef(lm_wetter_region)[1], 
           color = okabeito_colors()[2],
           size = 4,
           alpha = .5) +
  geom_label(aes(label = paste0("b", id),
                 x = grandmean + delta), 
             vjust = 1,
             size = 2,
             color = okabeito_colors()[1]) +
  geom_vline(xintercept = coef(lm_wetter_region)[1], 
             linetype = "dashed", color = okabeito_colors()[2]) +
  
  #coord_cartesian(xlim = c(7, 10), ylim = c(0, 16))  +
 
  annotate("label",
           y = 1,
           x = coef(lm_wetter_region)[1],
           vjust = 1,
           label = paste0("b0"),
           size = 2,
           color = okabeito_colors()[2]) +

  geom_segment(aes(yend = region, xend = temp), 
               x = coef(lm_wetter_region)[1], 
               color = okabeito_colors()[1])  +
  labs(y = "",
       x = "Temperatur")
```




# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


