---
title: Dead Man Bias in correlation
author: Sebastian Sauer
date: '2024-03-05'
slug: dead-man-bias-in-correlation
categories:
  - rstats
tags:
  - association
  - correlation
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
```


# Background

Nassim Taleb points out in the paper [Fooled by Correlation: Common Misinterpretations
in Social "Science"](https://datascienceassn.org/sites/default/files/Fooled%20by%20Correlation%20Common%20Misinterpretations%20in%20Social%20Science.pdf) that spurious correlation may appear due to various reasons.
One reason is what he calls the "Dead Man Bias", occuring if constant data is added to uniformly distributed data.
The thing is if the data is uniformly distributed and non-correlated, you will get a spurious correlation if constant data is added.


# Data example


Let's simulate a dataset, $n=10^4$ with $X,Y \sim Unif(0,1)$.

```{r}
set.seed(42)
n <- 1e4
d <-
  tibble(x = runif(n = n),
         y = runif(n = n))

d_dead_man <-
  tibble(x = rep(0, n * 0.2),
         y = x)

d2 <-
  d |> 
  bind_rows(d_dead_man)
```

Here's their correlation:

```{r}
d2 |> 
  summarise(cor_x_y = cor(x, y))
```


As can be seen, a *spuriou correlation* pops up.


```{r}
d2 |> 
ggplot( aes(x = x, y = y)) +
  stat_density_2d(aes(fill = ..density..), geom = "tile", contour = FALSE) +
  scale_fill_gradient(low = "blue", high = "red") +
  theme_minimal()
```



# Discussion

Note that this bias does not appear for normally distributed data.


# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


