---
title: Some ways to plot the distribution of each variable of a data frame
author: Sebastian Sauer
date: '2022-09-26'
slug: some-ways-to-plot-the-distribution-of-each-variable-of-a-data-frame
categories:
  - plotting
  - dataviz
  - rstats
tags:
  - plotting
  - ggplot2
  - tidyverse
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Motivation


Often times, in explorative data analysis, one would like to plot the distribution of the relevant variables.
Whereas ggplot provides handy tools to plot *one* variable after each other,
it would come handy to plot 'em all in one go.

Of course, there are many ways to achieve this comfort. This posts presents some of them in brevity.


# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
```



# Load data

Demo data.

```{r}
d_path <- "https://raw.githubusercontent.com/sebastiansauer/pradadata/master/data-raw/extra.csv"
d <- read_csv(d_path)
head(d)
```


Select the items (columns) we would like to plot the distribution for:


```{r}
d_items <- 
  d %>% 
  select(matches("i\\d\\d")) %>% 
  select(-i21)
```


```{r}
d_item_names <-
  names(d_items)

d_item_names
```


# Let's plot

## Way 1
```{r eval = TRUE}
map(d_item_names,
    ~ ggplot(data = d_items) + geom_bar(aes(x = .data[[.x]])) + labs(y = .x))
```


## Way 2

```{r eval = TRUE}
d_items %>% 
  pivot_longer(everything(), names_to = "item") %>%
  group_by(item) %>% 
  group_split() %>% 
  map(~ {ggplot(., aes(x = value)) + geom_bar()})
```


## Way 3


```{r fig.asp = 1}
d_items %>% 
  pivot_longer(everything(), names_to = "item") %>%
  mutate(item = factor(item, levels = d_item_names)) %>% 
  ggplot(aes(x = value)) +
  geom_bar() +
  facet_wrap(~ item)

```


## Way 4

```{r}
library(DataExplorer)

d_items %>% 
  plot_histogram()
```



# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


