---
title: Simple contingency tables in R
author: Sebastian Sauer
date: '2024-01-12'
slug: simple-contingency-tables-in-r
tags:
  - eda
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
```


# Motivation

Assume we would like to compute contingency tables in R without much ado. Let's explore some ways.

# Toy data

```{r}
data(mtcars)
```



# Using `table` and friends


```{r}
mtcars |> 
  select(vs, am) |> 
  table()
```

Let's add margins:

```{r}
mtcars |> 
  select(vs, am) |> 
  table() |> 
  addmargins()
```

Let's show rather proportions instead of raw counts:

```{r}
mtcars |> 
  select(vs, am) |> 
  table() |> 
  prop.table() |> 
  addmargins()
```

Note that `addmargins` must appear after `proptable`. 

Let's round to 2 places:


```{r}
mtcars |> 
  select(vs, am) |> 
  table() |> 
  prop.table() |> 
  addmargins() |> 
  round(2)
```


Let's change the margin, i.e, the groups which should sum up to 100%:


Proportions *per row*:

```{r}
mtcars |> 
  select(vs, am) |> 
  table() |> 
  prop.table(margin = 1) |> 
  addmargins() |> 
  round(2)
```

Proportion per *column*:

```{r}
mtcars |> 
  select(vs, am) |> 
  table() |> 
  prop.table(margin = 2) |> 
  addmargins() |> 
  round(2)
```



# Using `count`


```{r}
mtcars |> 
  count(vs, am)
```


Adding proportions:

```{r}
mtcars |> 
  count(vs, am) |> 
  mutate(prop = n/sum(n))
```


Since some while, `count` returns an ungrouped table, we would need to group according to our wishes.

If we group e.g., by `am` we get the proportion of `vs` values for each group of `am`:

```{r}
mtcars |> 
  count(am, vs) |> 
  group_by(am) |> 
  mutate(prop = n/sum(n))
```


Grouping by `vs`:

```{r}
mtcars |> 
  count(am, vs) |> 
  group_by(vs) |> 
  mutate(prop = n/sum(n))
```


We can also sum up all groups to 100%, if we do not group at all:


```{r}
mtcars |> 
  count(am, vs) |> 
  mutate(prop = n/sum(n))
```

# Don't drop unused factor levels

Use `count(..., .drop = FALSE)` to prevent that unused factor levels are dropped.




# See also

[This SO post](https://stackoverflow.com/questions/24576515/relative-frequencies-proportions-with-dplyr) shows a nice overview on grouping using `count` from `dplyr`.



# Conclusions


There are many ways in R to produce contingency tables. Here are two :-)



# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


