---
title: Checking Moodle test log data
author: Sebastian Sauer
date: '2022-02-08'
draft: false
slug: checking-moodle-test-log-data
categories:
  - rstats
tags:
  - teaching
editor_options: 
  chunk_output_type: console
---


# Motivation: Let's check whether server blackouts seem probable 

After one particular exam, a student complaint that Moodle was not reacting during some specified time period.

In this post, we'll check whether we find evidence in favor or against a failout of the server.


# Setup


```{r}
library("tidyverse")
library("digest")  # anonymize data
library("lubridate")  # working with dates/time
theme_set(theme_minimal())
```


# Anonymize data

Here's the path to the original data. You cannot access it, as it is confidential data.
In this step, I'd just like to show you how to anonymize it.

```{r results = "hide"}
d_path_raw <- "/Users/sebastiansaueruser/Downloads/logs-some-moodle-test.csv"

d <- read_csv(d_path_raw)
```

Here's the main take, we'll use the MD5 crypto function to get a fingerprint of the file:

```{r}
d2 <-
  d %>% 
  mutate(user = digest(`Vollständiger Name`, algo = "md5"),
         ip = digest("IP-Adresse", algo = "md5"))
```

Finally, let's prepocess a bit, ie. delete columns with sensitive data,
and parse the time:

```{r}
d3 <-
  d2 %>% 
  select(-c(2,3,7,9)) %>% 
  mutate(time_stamp = parse_datetime(Zeit, "%d.%m.%Y %H:%M")) %>% 
  select(-Zeit) %>% 
  mutate(id = row_number())
```


And now save the anonymized data:

```{r}
d_path_anonymized <- "/Users/sebastiansaueruser/github-repos/Lehre/data/d3.csv"

write_csv(d3, d_path_anonymized)
```


# Read anonymized data


```{r}
d_anom_path <- "https://raw.githubusercontent.com/sebastiansauer/Lehre/main/data/d3.csv"

d_anom <- read_csv(d_anom_path)
```



# Sense check

Let's see whether the time/dates appear sensible:

```{r}
d_anom %>% 
  summarise(min(time_stamp),
            max(time_stamp))
```


OK.


# Are there periods where no students interacted with Moodle?

If there were time periods were no student at all interacted with the Moodle server (according to the log data),
we'd have evidence for some potential Server failure.




# Preprocess data

Let's filter the log data for the exam time of the first exam (there were two on the relevant day):

```{r}
d_exam1 <- 
  d_anom %>%
  filter(time_stamp < ymd_hm("2022-01-31 09:20")) %>% 
  filter(time_stamp > ymd_hm("2022-01-31 07:55"))
```




Check for missing data:


```{r}
d_exam1 %>%
  summarise(across(everything(), ~ sum(is.na(.))))
```

OK, no missings, that's reassuring in terms of data quality.



# Overall times of server contact


```{r}
d_exam1 %>% 
  select(time_stamp) %>% 
  ggplot(aes(x = time_stamp)) +
  scale_x_datetime() +
  geom_histogram()
```


Hm...



```{r}
d_exam1 %>% 
  select(time_stamp) %>% 
  ggplot(aes(x = time_stamp)) +
  scale_x_datetime() +
  geom_density()
```


Looks unsuspicious.


Let's depict each server interaction as a dot:

```{r}
d_exam1 %>% 
  cbind(y = runif(n = nrow(d_exam1))) %>% 
  relocate(y, .before = 1) %>% 
  relocate(time_stamp, .before = 2) %>% 
  ggplot(aes(x = time_stamp, y = y)) +
  geom_point(alpha = .1) +
  scale_x_datetime() +
  scale_y_continuous(limits = c(0, 1))
```



Let's zoom in right before the end of the test:



```{r}
d_exam1 %>% 
  cbind(y = runif(n = nrow(d_exam1))) %>% 
  relocate(y, .before = 1) %>% 
  relocate(time_stamp, .before = 2) %>% 
  ggplot(aes(x = time_stamp, y = y)) +
  geom_point(alpha = .1) +
  scale_x_datetime(limits = c(ymd_hm("2022-01-31 08:50"), ymd_hm("2022-01-31 09:00"))) +
  scale_y_continuous(limits = c(0, 1))
```



It appears the time wise resolution of our data does not allow for finer grains than 1 one minute windows.



Let's try something more quantitatively looking:



```{r}
d_exam1 %>% 
  select(time_stamp) %>% 
  filter(between(time_stamp, ymd_hm("2022-01-31 08:00"), ymd_hm("2022-01-31 08:59"))) %>% 
  count(time_stamp) %>% 
  ggplot() +
  geom_col(aes(x = time_stamp, y = n))
```


Looks unsuspicious.



# Debrief

Not finding evidence of some failure is not evidence of no failure, in general at least.


