---
title: Data sets for for teaching
author: Sebastian Sauer
date: '2022-02-23'
slug: data-sets-for-for-teaching
categories:
  - rstats
tags:
  - data
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
```



# Data

Here's a opionated list of data sets useful for teaching purposes:


- `mtcars`
  - [csv](https://vincentarelbundock.github.io/Rdatasets/csv/datasets/mtcars.csv)
  - [doc](https://vincentarelbundock.github.io/Rdatasets/doc/datasets/mtcars.html)
- `tips`
  - [csv](https://vincentarelbundock.github.io/Rdatasets/csv/reshape2/tips.csv)
  - [doc](https://vincentarelbundock.github.io/Rdatasets/doc/reshape2/tips.html)
- `flights` (NYCflights13)
  - [csv](https://vincentarelbundock.github.io/Rdatasets/csv/nycflights13/flights.csv)
  - [doc](https://vincentarelbundock.github.io/Rdatasets/doc/nycflights13/flights.html)
- `Saratoga houses`
  - [csv](https://vincentarelbundock.github.io/Rdatasets/csv/mosaicData/SaratogaHouses.csv)
  - [doc](https://vincentarelbundock.github.io/Rdatasets/doc/mosaicData/SaratogaHouses.html)
- `diamonds`
  - [csv](https://vincentarelbundock.github.io/Rdatasets/csv/ggplot2/diamonds.csv)
  - [doc](https://vincentarelbundock.github.io/Rdatasets/csv/ggplot2/diamonds.html)
- `wo_men`
  - [csv](https://raw.githubusercontent.com/sebastiansauer/pradadata/master/data-raw/wo_men.csv)
- `OECD well-being`
  - [source](https://data-se.netlify.app/2018/10/16/oecd-wellbeing-dataset-2016/)
- `penguins`
  - [csv](https://vincentarelbundock.github.io/Rdatasets/csv/palmerpenguins/penguins.csv)
  - [doc](https://vincentarelbundock.github.io/Rdatasets/csv/palmerpenguins/penguins.html)
- `Ames housing`
  - [source on Kaggle, login needed](https://www.kaggle.com/prevek18/ames-housing-dataset)
- `teacher rating`
  - [csv](https://vincentarelbundock.github.io/Rdatasets/doc/AER/TeachingRatings.csv)
  - [doc](https://vincentarelbundock.github.io/Rdatasets/doc/AER/TeachingRatings.html)


Note that the data sets are provided as standard CSV files (comma separeted, dots as delimiters).



# Data repositories

Check this out for longish compilations of data sets useful for data analysis:


- [Kaggle](https://www.kaggle.com/datasets)
- [R Datasets](https://vincentarelbundock.github.io/Rdatasets/articles/data.html)


# How to import into R


1. Open the web page of the data by clicking on the respective `[csv]` link (above)
2. Copy the URL of that web page to your clipboard (control-c)
3. Paste the URL to the following R code
4. Execute the R code below:


```{r}
# replace the following URL with your URL
data_url <- "https://vincentarelbundock.github.io/Rdatasets/csv/AER/TeachingRatings.csv"

library(tidyverse)  # remember to install once upfront

d <- read_csv(data_url)
```

Check it:


```{r}
glimpse(d)
```

Yeah!

Don't worry about the first column with the strange name `...1`, that's
nothing but a technical artifact. Just ignore it.



# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


