---
title: Consistency of set.seed across different systems
author: Sebastian Sauer
date: '2022-12-13'
slug: consistency-of-set-seed-across-different-systems
categories:
  - rstats
tags:
  - simulation
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
library(digest)
library(rstanarm)
```


# Motivation

Reproducibility of results is a major concern in science and industry alike.
However, there are numerous pitfalls which may threaten reproducibility.
This post explores one possible issue - the fixation of start values of random numbers drawn by R.

Reasons that seeds in R across different systems may (possibly) vary (I don't know if they do, but I heard so) may include:

- different OSes
- different version of OSes
- different versions of R


# User error

Maybe the most straight forward explanation is a user error.
Don't forget to call `set.seed()` right  before the calculation.
Be sure to use a pre defined seed value.



# Your help needed

Please let me know whether (or not) you get the same random numbers,
once the seed is fixed.
In order to check the possible influence source,
please note your OS, and the R version (or paste the session info).

Thanks! 




# Same random numbers 

What `set.seed()` does is the fix the value of `.Random.seed`,
which is in turn used to initiate the random number generator of R.

```{r}
.Random.seed %>% head()
```



## Without seed

Without seed, the random numbers drawn by R will be transient, ie., subject of change:

```{r}
rnorm(1)
```


```{r}
replicate(n = 10, expr = rnorm(1))
```


## With seed


```{r rnorm1-with-seed}
set.seed(42)

rnorm(1)
```





```{r}
set.seed(42)
replicate(n = 10, expr = rnorm(1))
```


## Using a hash 

Let's use a hash number to control the output more succinctly:


```{r}
set.seed(42)
x <- replicate(n = 10, expr = rnorm(1))

digest(x)
```


Let's recompute and compare the hash values:



```{r}
set.seed(42)
x2 <- replicate(n = 10, expr = rnorm(1))

digest(x2)
```



```{r}
identical(x, x2)
```

Identical, as expected.




# Seeds in regression models


## lm

```{r}
set.seed(42)
lm1 <- lm(mpg ~ hp, data = mtcars)
lm1_coef <- coef(lm1)
lm1_coef
```


Rerun to double check:

```{r}
set.seed(42)
lm1 <- lm(mpg ~ hp, data = mtcars)
lm1_coef <- coef(lm1)
lm1_coef
```

## Stan mtcars


```{r stan1}
lm2 <- stan_glm(mpg ~ hp, data = mtcars, seed = 42, refresh = 0)
lm2_coef <- coef(lm2)
lm2_coef
```


Rerun:



```{r stan2}
lm2 <- stan_glm(mpg ~ hp, data = mtcars, seed = 42, refresh = 0)
lm2_coef <- coef(lm2)
lm2_coef
```


## Stan penguins

Import data and run the model:




```{r}
penguins <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/palmerpenguins/penguins.csv")

library(rstanarm)
lm3 <- stan_glm(body_mass_g ~ bill_length_mm, 
                seed = 42,
                refresh = 0,
                data = penguins)
coef(lm3)
```


By the way: 

Copy-paste the chunk above (which is self-containing in terms of data and R packages) and then type `reprex::reprex()` in the console (you need to have the package  `reprex` installed). You'll get a reproducible example on your clipboard.


Rerun:


```{r}
lm3 <- stan_glm(body_mass_g ~ bill_length_mm, 
                seed = 42,
                refresh = 0,
                data = penguins)
coef(lm3)
```


# Session info

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()
```




