---
title: 'Exercises (no solutions): data vizualization on flight delays using tidyverse
  tools'
author: Sebastian Sauer
date: '2021-02-24'
slug: exercises-no-solutions-data-vizualization-on-flight-delays-using-tidyverse-tools
categories:
  
tags:
  - tutorial
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
```


# Get the data

We'll be analyzing the data set `flights`, describing the flights which started from NYC in 2013.

Here's how to get the data set:


```{r message = FALSE}
library(tidyverse)
library(nycflights13)

data("flights")
```


## Alternative way to get the data

Alternatively, import the data from a csv file:

```{r}
flights_url <- "https://vincentarelbundock.github.io/Rdatasets/csv/nycflights13/flights.csv"
flights <- read_csv(flights_url)
```


## Code book

Find the code book of the data set [here](https://vincentarelbundock.github.io/Rdatasets/doc/nycflights13/flights.html).

Alternatively, try `? nycflights13`. 



# Exercises

1. Plot the distribution of the delays. Describe your insights.

2. Plot the distribution of the delay per origin airport.

3. Visualize the assocation of delay and time of the day. Find a way to reduce overplotting. Hint: Try out `geom_bind2d()` instead of using a scatter plot.

4. Visualize the assocation of delay and distance to destination. Separate by origin and month.

5. Visualize the assocation of delay and time of the day. Only include the three airlines where the delay is highest. Reduce overplotting.

6. Visualize the proportion of delayed flights per origin.

7. Visualize the proportion of delayed flights per time.

8. Visualize the proportion of delayed flights per week day.



# Solutions


Check out the solutions [here](https://data-se.netlify.app/2021/02/24/case-study-data-vizualization-on-flight-delays-using-tidyverse-tools/). Only after trying really hard 😀.



# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()


