---
title: Rowwise NA
author: Sebastian Sauer
date: '2022-05-09'
slug: rowwise-na
categories:
  - rstats
tags:
  - tutorial
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
```


# Sample data


```{r}
data("mtcars")
```


Create some NA:


```{r}
mtcars$mpg[c(1,2,3)] <- NA
mtcars$hp[c(1,2,3)] <- NA
```


# Count NA rowwise


What we would like to achieve is to comfortable count the missing values *per row*.

Define helper function:

```{r}
sum_isna <- function(x) sum(is.na(x))
```


# Way 1: rowwise sum with mutate and `c_across`

```{r}
mtcars %>% 
  rowwise() %>% 
  mutate(Na_n = sum_isna(c_across(everything()))) %>% 
  ungroup()

```


A more in-depth treatment can be found [here](https://dplyr.tidyverse.org/articles/rowwise.html).



# Way 2: `apply()` with margin 1

Margin 1 means rowwise:

```{r}
mtcars %>% 
  mutate(Na_n = apply(mtcars, 1, sum_isna))
```



# Way 3: `rowSums`


```{r}
mtcars %>% 
  mutate(Na_n = rowSums(is.na(mtcars)))
```


[Source](https://stackoverflow.com/questions/64356552/count-missing-values-with-rowwise-and-add-number-of-missing-values)


# Way 4: `cur_data()`


`cur_data()` refers to the data in the *current* group, which is, in the case
of `rowwise()` the current row.

```{r}
mtcars %>% 
  rowwise() %>% 
  mutate(NA_n = sum_isna(cur_data()))
```

# Why not `map()`?


When using `map()` in relation with `mutate()`, we map a *column* of the data frame to some function.
However, when counting missing values per row, we would like to map a row to a function,
which is not possible using `map()`. 




# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


