---
title: German weather
author: Sebastian Sauer
date: '2022-07-24'
slug: german-weather
categories:
  - rstats
tags:
  - dataviz
  - datawrangling
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
library(easystats)
library(ggplot2); theme_set(theme_minimal())  # ggplot theme
```

# Motivation

Let's explore the change over time in German weather.
We are not pretending doing real meteorology here; rather, we are playing around a bit.
At the very least, we'll do over own analyses, so we know what's going in.
Data stems from [Deutscher Wetter Diensts, DWD](https://www.dwd.de/DE/Home/home_node.html).



# Load data

The DWD data has been prepared from fast consumption as [laid out in this post](https://data-se.netlify.app/2022/07/24/preparing-german-weather-data/).

As a result, the data have been made available [in this repo](https://github.com/sebastiansauer/datasets).

```{r read-data}
d_path <- "https://raw.githubusercontent.com/sebastiansauer/datasets/main/csv/precip_temp_DWD.csv"

d <- read_csv(d_path)
```


```{r}
glimpse(d)
```




# Main temperature trajectory over time

## Visualization

```{r vis1}
d_grouped <-
  d %>% 
  group_by(region, year) %>% 
  summarise(temp_avg = mean(temp))

d_grouped %>% 
  ggplot(aes(x = year, y = temp_avg, group = region)) +
  geom_line(color = "grey60") +
  labs(title = "Temperature change over time in Germany",
       caption = "Data Source: DWD",
       subtitle = "One grey line depicts one Bundesland") +
  geom_smooth(method = "lm", group = 1) 
```

Temperature is on the rise, it appears.

## Linear model


```{r m1}
m1 <- lm(temp_avg ~ year, data = d_grouped)
print_html(model_parameters(m1))
```

This amounts to approx. 0.1 degrees Celcius per Dekade on average.


```{r}
plot(parameters(m1), show_intercept = TRUE)
```

# Temperature change per month

## Vis 1: Change per Month for whole of Germany

```{r  vis2, fig.asp = 1.5}
d_grouped2 <-
  d %>% 
  group_by(region, year, month) %>% 
  summarise(temp_avg = mean(temp))

d_grouped2 %>% 
  ggplot(aes(x = year, y = temp_avg, group = region)) +
  geom_line(color = "grey60") +
  labs(title = "Temperature change over time per month") +
  facet_wrap(~ month, ncol = 3) +
  geom_smooth(method = "lm", group = 1)
```
It looks like that for each month, temperature is on the rise.

I don't think it is very important to check this trend for each Bundesland. 
However, it may well be that this trend is more pronounced for some Bundesländer as opposed to others.




## Linear model

```{r m2}
m2 <- lm(temp_avg ~ as.factor(month) + year, data = d_grouped2)
print_html(model_parameters(m2))
```



```{r}
plot(parameters(m2), show_intercept = TRUE)
```


Here we see the effect of the month, with the Month of July being the warmest, which makes sense.
Note that the intercept is not interpretable here as it shows the temperature in year 0, which is a bit far away.

For comparison, let's just see the avarage temperatures per month:

```{r m3}
m3 <- lm(temp_avg ~ as.factor(month), data = d_grouped2)
print_html(model_parameters(m3))
```

Note that these are the averages across all the years in the data set.


```{r}
plot(parameters(m3), show_intercept = TRUE)
```


## Vis 2: Trend by Bundesland


```{r fig.asp = 1.5}
d_grouped2a <-
  d %>% 
  group_by(region, year) %>% 
  summarise(temp_avg = mean(temp))

d_grouped2a %>% 
  ggplot(aes(x = year, y = temp_avg)) +
  geom_line(color = "grey60") +
  labs(title = "Temperature change over time per Bundesland") +
  facet_wrap(~ region, ncol = 3) +
  geom_smooth(method = "lm")
```

```{r}
m4 <- lm(temp ~ region + year, data = d)
print_html(parameters(m4))
```

The point estimate indicates that Saarland gained most in terms of temperature;
however, the differences are not clear cut between the Bundesländer,
so no strong conclusion can be drawn.

# Change per decade


```{r}
d2 <- 
  d %>% 
  mutate(decade = trunc(year  / 10))

head(d2)
```


```{r}
d_grouped3 <-
  d2 %>% 
  group_by(region, decade) %>% 
  summarise(temp_avg = mean(temp)) %>% 
  mutate(temp_dec_before = lag(temp_avg)) %>% 
  mutate(temp_change_decade = temp_avg - temp_dec_before)

d_grouped3 %>% 
  head()
```


## Vis 1

```{r}
d_grouped3 %>% 
  ggplot(aes(x = decade, y = temp_avg)) +
  geom_point(color = "grey60") +
  labs(title = "Temperature change over decades in Germany",
       caption = "Data Source: DWD",
       subtitle = "One grey dot depicts one Bundesland",
       y = "Temperature in °C") +
  stat_summary(geom = "line", fun = mean) +
  scale_x_continuous(breaks = 188:202)
```

Again, we see that temperature is on the rise.

## Vis 2: Temperature change per decade

```{r}
d_grouped3 %>% 
  ggplot(aes(x = decade, y = temp_change_decade)) +
  geom_point(color = "grey60") +
  labs(title = "Temperature change over decades in Germany",
       caption = "Data Source: DWD",
       subtitle = "One grey dot depicts one Bundesland") +
  stat_summary(geom = "line", fun = mean) +
  scale_x_continuous(breaks = 188:202)
```

For any to given subsequent decades, we see that the the average temperature of the decase has risen.
This trend is most pronounced in the 1990 and 2000 decades;
but it is positive (hotter) for any data point in this graph *AFTER* 1960.


# Change in variability


```{r}
d_grouped4 <-
  d2 %>% 
  group_by(region, decade) %>% 
  summarise(temp_sd = sd(temp)) %>% 
  mutate(temp_sd_dec_before = lag(temp_sd)) %>% 
  mutate(temp_sd_change_decade = temp_sd - temp_sd_dec_before)

d_grouped4 %>% 
  head()
```



## Vis 

```{r}
d_grouped4 %>% 
  ggplot(aes(x = decade, y = temp_sd)) +
  geom_point(color = "grey60") +
  labs(title = "Temperature *variability* change over decades in Germany",
       caption = "Data Source: DWD",
       subtitle = "One grey dot depicts one Bundesland") +
  stat_summary(geom = "line", fun = mean) +
  scale_x_continuous(breaks = 188:202)
```


Apart from the 1940, there's no apparent change in variability (SD) per decade.



# Debrief


This post is NOT real meteorology science!
Rather, it is some exploratory appraoching to a highly sophisticated field of research. Certainly,
the state of art is not reflected in this post.
The (aspired) benefit of this post is to bring the data-savy layman closer to weather data in an attempt not to have to (completely) rely on third persons.


# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


