---
title: Comparing Jamovi and rstanarm
author: Sebastian Sauer
date: '2022-05-09'
slug: comparing-jamovi-and-rstanarm
categories:
  - rstats
tags:
  - Bayes
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
```


# Motivation

Let's try to see how much the results of Jamovi (2.2.5) and rstanarm (2.21.1) converge.
It's probably difficult to say because the defaults are different,
and it may not be straight forward to translate back and forth.

Anyhow, let's see.


# data

The ubiqious `mtcars`.


```{r}
data(mtcars)

write_csv(mtcars, file = "mtcars.csv")
```

# Model 1

Let's try: `mpg ~ am`.

## rstanarm

```{r}
library(rstanarm)

m1 <- stan_glm(mpg ~ am, data = mtcars, refresh = 0)

posterior_interval(m1)
```

I did not touch the defaults, which can be accessed like this:

```{r}
prior_summary(m1)
```


## Jamovi

I did not touch the defaults, which is a JZS prior with scale r = 0.354.

Jamovi report a 95% CI of [3.07, 9.88] which is quite close to what rstanarm found.


# Model 2


`mpg ~ am + hp`


## rstanarm


```{r}
m2 <- stan_glm(mpg ~ am + hp, data = mtcars, refresh = 0)
posterior_interval(m2)
```



## Jamovi


Jamovi provides R code for Bayes regresison, but it not yet kind of released.


```{r eval = FALSE}
jsq::blinReg(
    data = data,
    dep = mpg,
    covs = vars(am, hp),
    postSumm = TRUE,
    modelTerms = list(
        list(
            components="am",
            isNuisance=FALSE),
        list(
            components="hp",
            isNuisance=FALSE)))
```


The source code can be accessed [here](https://github.com/jonathon-love/jsq) though.

Anyhow, it produces (in Jamovi) this output:

```
#
#  Model Comparison                                                            
#  ─────────────────────────────────────────────────────────────────────────── 
#    Models        P(M)      P(M|data)    BF-M           BF-10        R²       
#  ─────────────────────────────────────────────────────────────────────────── 
#    Null model    0.2500     2.919e-8       8.756e-8        1.000    0.0000   
#    am + hp       0.2500     0.998333    1796.888017     3.421e+7    0.7820   
#    hp            0.2500     0.001664       0.005001    57020.790    0.6024   
#    am            0.2500     2.549e-6       7.648e-6       87.348    0.3598   
#  ─────────────────────────────────────────────────────────────────────────── 
# 
#
#
#
#  Posterior Summaries of Coefficients                                                                        
#  ────────────────────────────────────────────────────────────────────────────────────────────────────────── 
#    Coefficient    Mean        SD          P(incl)    P(incl|data)    BF-inclusion    Lower       Upper      
#  ────────────────────────────────────────────────────────────────────────────────────────────────────────── 
#    Intercept      20.09062    0.514278     1.0000          1.0000           1.000    19.04175    21.13950   
#    am              5.11035    1.062349     0.5000          0.9983         599.882     2.94368     7.27703   
#    hp             -0.05703    0.007732     0.5000          1.0000      387822.393    -0.07280    -0.04126   
#  ────────────────────────────────────────────────────────────────────────────────────────────────────────── 
# 
# 
#
```


Which is in turn quite comparable to rstanarm.


## Interim conclusion

This little exercise shows that the two software packages *may* yield similar results,
at least in some cases.





# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


