---
title: Setting to NA, conditionally
author: Sebastian Sauer
date: '2023-01-16'
slug: setting-to-na-conditionally
categories:
  - rstats
tags:
  - tidyverse
  - datawrangling
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
```


# Motivation

Let's assume we would like to change the values of multiple variables depending in the state of another variable.
For the sake of concreteness, let's say we have some variable called `data_trustworthiness`. If this variable (indicating whether or not we can have confidence in some other variables) has the value `FALSE` for some cases, we would like to set the varialbe `measure1` and `measure2` to `NA`, thus reflecting that the data from our measurements are not reliable.

Let's use the tidyverse ecosystem.

# Minimal example

For the sake of simplicity, we'll make use of the `mtcars` dataset.

First, we build a variable `trustworthy` with (for this minimal example) random values.

```{r}
data(mtcars)

mtcars <-
  mtcars |> 
  mutate(trustworthy = sample(c(1, 0), size = nrow(mtcars), replace = TRUE)) |> 
  relocate(trustworthy)

glimpse(mtcars)  # check
```


Next comes the actual *conditional change* of the data variables (`data_vars`).

```{r}
mtcars_new <- 
  mtcars |> 
  mutate(across(.cols = c("mpg", "cyl"),
                .fns = ~ifelse(trustworthy == 0, NA, .x))) 

glimpse(mtcars_new)  # check
```

```{r}
sum(is.na(mtcars_new))
```


```{r}
head(mtcars_new)
```


# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


