---
title: Working with list columns - an example
author: Sebastian Sauer
date: '2024-10-13'
slug: working-with-list-columns-an-example
categories:
  - rstats
tags:
  - loops
  - purrr
  - list
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
```


# Introduction

In this post, I want to show you how to work with list columns in R. List columns are a powerful feature of the `tidyverse` that allow you to store multiple objects in a single column of a data frame. This can be useful when you have a list of objects that you want to keep together, such as a list of data frames or a list of models.


# Example data

```{r}
my_path <- paste0(here::here(),"/content/post/2024-10-13-working-with-list-columns-an-example/df_with_list_columns.rds")

file.exists(my_path)

df <- read_rds(my_path)
```


# Add list column 1

```{r}
df |>  
  mutate(id_is_seq_1_to_n = map_lgl(data, function(df) identical(df$id, 1:88)))
```




# Add list column 2

```{r}
df |>  
  mutate(id_is_seq_1_to_n = map_lgl(data, ~ identical(.x$id, 1:88)))
```


# Extract list column

If we pull the list columns `data`, we will get a list of data frames:

```{r}
df |> 
  pull(data) |> 
  str()
```

Same thing using the dollar operator:

```{r}
df$data |> 
  str()
```

Or double brackets:

```{r}
df[["data"]] |> 
  str()
```


Pulling only the n-th element can be achieved like this:

```{r}
df$data[[1]] |> 
  str()
```



# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


