---
title: 'Trying tidymodels: step_num2factor'
author: Sebastian Sauer
date: '2020-12-12'
slug: trying-tidymodels-step-num2factor
categories:
  
tags:
  - prediction
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
library(tidymodels)  # modelling
```


# Understanding recipes and preprocessing


Having defined a recipe in [this Kaggle competition](https://www.kaggle.com/ssauer/tidymodels-titanic/report), I was left wondering about some details of the recipe definition. Let's explore that.


# Cross-posting note

This issue has been place on [SO here]()




```{r}
data <- 
  tibble::tribble(
  ~Survived, ~Pclass,
         1L,      1L,
         1L,      2L,
         0L,      3L
  )

```


```{r}
data$Survived <- as.numeric(data$Survived)
```




# Define recipe

```{r}
titanic_recipe <- 
  
  # define model formula:
  recipe(Survived ~ Pclass, data = data) %>%
  
  # convert numeric outcome to nominal (factor):
  step_num2factor(Survived,
                  levels = c("dead", "alive"))   
  
```




# Prepare (`prep()`) the recipe


```{ error = TRUE}
prep(titanic_recipe) 
```


What's going on here? Why the error? 


# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()


