---
title: Prevent dropping from non-occuring levels using dplyr
author: Sebastian Sauer
date: '2024-01-30'
slug: prevent-dropping-from-non-occuring-levels-using-dplyr
categories:
  - rstats
tags:
  - tidyverse
  - datawrangling
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
```


# Problem

Consider the following situation:


```{r}
mtcars |> 
  group_by(high_hp = hp > 1000) |> 
  count(high_hp)
```

The summary table does not show the level `TRUE`, as it is not occuring in the data.
This can be problematic: If the data is unknown before summarizing and you would expect that both/all levels (TRUE, FALSE) occur. Just imagine that a subsequent function will count the level `TRUE` and the level `FALSE`. If one level is missing, your system may brake down.


# Solution

Using dplyr and `count` or `summarize`, one solution is to turn the grouping variable in a factor,
and then using the parameter `.drop = FALSE` in `count` or `summarise`. 

Here's an example for `count`:


```{r}
mtcars |> 
  group_by(high_hp = hp > 1000) |> 
  mutate(high_hp = factor(high_hp, levels = c(FALSE, TRUE))) |> 
  count(high_hp, .drop = FALSE)
```

And here a similar one for `summarise`:

```{r}
mtcars %>% 
  group_by(hp_over_1000 = factor(hp > 1000, levels = c(FALSE, TRUE)), .drop = FALSE) %>% 
  summarise(mean_hp = mean(hp))
```

If we would like to make sure that the resulting variable is not `NaN`, then we could use `complete`:

```{r}
mtcars %>% 
  group_by(hp_over_1000 = factor(hp > 1000, levels = c(FALSE, TRUE)), .drop = FALSE) %>% 
  summarise(mean_hp = mean(hp)) %>%
  complete(hp_over_1000, fill = list(mean_hp = NA))
```




# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


