---
title: Penguins Lasso with Tidymodels
author: Sebastian Sauer
date: '2022-07-24'
slug: penguins-lasso-with-tidymodels
categories:
  - rstats
tags:
  - prediction
  - tidymodels
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```






# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
library(tidymodels)
```


# Data

```{r}
data("penguins", package = "modeldata")
```



# A bit more than minimal


```{r}
rec1 <-
  recipe(body_mass_g ~ ., data = penguins) %>% 
  step_dummy(all_nominal()) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_nzv(all_numeric_predictors()) %>% 
  step_naomit(all_predictors())
```


Checks:

```{r}
summary(rec1)
```


```{r}
tidy(rec1)
```


```{r}
rec1 %>% 
  prep() %>% 
  bake(new_data = NULL) %>% 
  head()
```



```{r}
wf1 <-
  workflow() %>% 
  add_model(spec = linear_reg(engine = "glmnet", mixture = 1, penalty = 1)) %>% 
  add_recipe(recipe = rec1)
```

```{r}
m2 <- wf1 %>% 
  fit(data = penguins)
```


# Results

Use `tidy()` to extract model coefficients:



```{r}
tidy(m2)
```


In case some predictor beta has been shrunken to zero, it would be noted here.

# Extract fit

Less einformative ...


```{r}
m2 %>% 
  extract_fit_parsnip()
```

Similar to:    

```{r eval = TRUE}
m2 %>% 
  extract_fit_engine()
```


More information, such as the difference between the last two functions, can be found [here](https://www.tidymodels.org/learn/models/coefficients/), the tidymodels page on model coefficients.



# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


