---
title: Plot timelines using ggplot
author: Sebastian Sauer
date: '2022-11-30'
slug: plot-timelines-using-ggplot
categories:
  - rstats
tags:
  - visualization
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
library(vistime)  # time line
```



# Motivation

For project planing,
a visualization of some time line is often useful. 
If it's not the dates but rather the steps of a process, 
a graph of steps is more appropriate.
However, if the sequence of steps is simple and rather linear,
and the dates are the important piece of information to be transmitted,
a kind of *timeline* graph is warranted.

Here's one go: We're going to make use of Sandra Raabe's nice R package [vistime](https://shosaco.github.io/vistime/).


# Sample data

```{r}
timeline_data <- data.frame(
  event = c("Anmeldung", "Vergabe", "Schreiben", "Thesismodul", "Beginn Semester 1", "Ende Semester 1", "Beginn Semester 2", "Ende Semester 2"),
  start = c("2023-03-15", "2023-05-22", "2023-10-04", "2023-10-04", "2023-03-15", "2023-06-30", "2023-10-04", "2024-01-14"), 
  end   = c("2023-05-21", "2023-06-30", "2024-01-14", "2024-01-14", "2023-03-15", "2023-06-30", "2023-10-04", "2024-01-14"),
  group = c("Fachsemester 6", "Fachsemester 6", "Fachsemester 7", "Fachsemester 7", "Fachsemester 7", "Fachsemester 6", "Fachsemester 7", "Fachsemester 7")
)
```


# Visualization


```{r}

  
  gg_vistime(timeline_data, optimize_y = TRUE) +
  labs(title = "Thesis schreiben im Studiengang AWM",
       subtitle = "ENTWURF") +
  scale_x_continuous(
     breaks = c(as.numeric(as.Date("2023-03-01")),
               as.numeric(as.Date("2023-06-30"))))
    
    breaks = c(1.68e09, 1.685e9),
    labels = c("kjl", "jk"))
    
    breaks = c(as.numeric(as.Date("2023-03-01")),
               as.numeric(as.Date("2023-06-30"))),
    labels = c("jk", "§j"))


  scale_x_continuous(labels = c("Beginn FS 6", "Ende FS 6", "Beginn FS 7", "Ende FS 7"))


```


# Debrief



# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


