---
title: Differences according to importing CSV using different functions
author: Sebastian Sauer
date: '2023-01-19'
slug: differences-according-to-importing-csv-using-different-functions
categories:
  - rstats
tags:
  - datawrangling
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
library(easystats)
library(digest)  # hashes
```


# Motivation

Importing a CSV file can yield to - slightly - different results, according to which functions are used for importing the file. 
The question is whether the *data* itself is constant across different methods,
which is a neccessary condition for reliable analysis,
or at least the importing function must be known for a reproducible analysis,
in case different data can result when different import functions are used.


In this post, we will examine the effect of importing data using different functions.



# Data

We'll use the `penguins` data set.

```{r}
data_url <- "https://vincentarelbundock.github.io/Rdatasets/csv/palmerpenguins/penguins.csv"
```


# Method 1: `read.csv`

`read.csv` is a function from Base R.

Let's try it.

```{r}
d1 <- read.csv(data_url)
head(d1)
```


# Method 2: `read_csv`

```{r}
d2 <- read_csv(data_url)
head(d2)
```


# Method 3: `data_read`

```{r}
d3 <- data_read(data_url)
head(d3)
```

# First glimpse


```{r}
glimpse(d1)
glimpse(d2)
glimpse(d3)
```

Nothing that peeks into the eye.

# Hashes

A hash is like a fingerprint of a digital object - it is (quasi) unique.
Let's compute the hashes of the data sets. 
Note that we should preclude the first column as it's name is set differently by the function.

```{r}
d1 <- d1 %>% select(-1)
d2 <- d2 %>% select(-1)
d3 <- d3 %>% select(-1)
```


To get the hash value of some objects, we can use the function `digest()`.

```{r}
d1_hash <- 
d1 %>% 
  digest()

d1_hash
```


```{r}
d2_hash <- 
d2 %>% 
  digest()

d2_hash
```


```{r}
d3_hash <- 
d3 %>% 
  digest()

d3_hash
```


# Not exactly identical

As the hashes (fingerprints) differ, we can conclude that the objects are not *exactly* identical.
However, the differences may stem from subtle variations such as atttributes or classes of the data frame.

Let's focus on the data instead.


# Data comparison

By formatting as a `matrix`, we get rid of possible overheads of dataframes, leaving the pure data.

```{r}
d1_matrix <-
  d1 %>% 
  as.matrix()

d2_matrix <-
  d2 %>% 
  as.matrix()

d3_matrix <-
  d3 %>% 
  as.matrix()
```


Let's checkt the attributes of the matrices:

```{r}
d1_matrix %>% attributes()
d2_matrix %>% attributes()
d3_matrix %>% attributes()
```

Identical.

Now let's check the hashes of the matrices.

```{r}
d1_matrix_hash <- d1_matrix %>% digest()
d2_matrix_hash <- d2_matrix %>% digest()
d3_matrix_hash <- d3_matrix %>% digest()
```


```{r}
d1_matrix_hash
d2_matrix_hash
d3_matrix_hash
```

Identical.

# Conclusion

We can conclude that the data is identical across the methods (leaving the first column aside).

Note that no random numbers where involved in this analysis.



# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


