---
title: Predicting survival (Titanic) using a tidymodels approach
author: Sebastian Sauer
date: '2020-12-08'
draft: TRUE
slug: predicting-survival-titanic-using-a-tidymodels-approach
categories:
  - rstats
tags:
  - prediction
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```


# Motivation

Let's predict survival at the [Titanic disaster](https://www.kaggle.com/c/titanic) using the Kaggle competition data, download from [here](https://www.kaggle.com/c/titanic/data).

We'll make use of the [tidymodels](https://www.tidymodels.org/start/models/) approach.



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
library(tidymodels)  # modelling
library(broom)  # tidy model output
```




# Load data



```{r}
traindata_url <- "https://raw.githubusercontent.com/sebastiansauer/Statistiklehre/main/data/titanic/train.csv"
testdata_url <- "https://raw.githubusercontent.com/sebastiansauer/Statistiklehre/main/data/titanic/test.csv" 
```




```{r}
train <- read_csv(traindata_url)
test <- read_csv(testdata_url)
```



```{r}
glimpse(train)
```



# Data preprocessing


```{r}
train <-
  train %>% 
  mutate(Survived = factor(Survived))
```





# Define recipe

```{r}
titanic_recipe <- 
  recipe(Survived ~ ., data = train) %>% 
  update_role(Name, PassengerId,  new_role = "ID") %>% 
  step_rm(Ticket, Cabin, Name, PassengerId) %>% 
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_predictors()) 
```



```{r}
titanic_recipe
```



# Define model

```{r}
lr_mod <- 
  logistic_reg() %>% 
  set_engine("glm") %>% 
  set_mode("classification")
```


```{r}
lr_mod
```



# Define modelling workflow


```{r}
titanic_wf1 <-
  workflow() %>% 
  add_model(lr_mod) %>% 
  add_recipe(titanic_recipe)
```


```{r}
titanic_wf1
```



# Fit model


```{r}
titanic_fit_lr <-
  titanic_wf1 %>% 
  fit(data = train)
```


Model performance on *train* data:


```{r}
titanic_fit_lr %>% 
  pull_workflow_fit() %>% 
  tidy()
```

```{r}
titanic_fit_lr %>% 
  pull_workflow_fit() %>% 
  glance()
```





# Predict


```{r}
test <-
  test %>% 
  bind_cols(predict(titanic_fit_lr, test, type = "class"))
```




# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()


