---
title: FontAwesome in R and R Markdown
author: Sebastian Sauer
date: '2022-07-27'
slug: fontawesome-in-r-and-r-markdown
categories:
  - rstats
tags:
  - visualization
  - RMarkdown
  - markdown
output:
  blogdown::html_page:
    toc: yes
    number_sections: yes
---



```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "100%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)
```



# Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(tidyverse)  # data wrangling
```


# Use Case


Some times some nice emojis or icons are of benefit for your new post, right?

But, what's a useful way to implement icons?


# Way

Here's a quick way of incorporating [FontAwesome](https://fontawesome.com/) icons to your RMarkdown document:


````markdown

<center>
<font size="15"> 
```{r fa, message=FALSE, echo = TRUE, results='hold', eval=FALSE}
library(fontawesome)
fa("r-project", fill = "steelblue")
fa("bolt-lightning", fill = "steelblue")
fa("discourse", fill = "steelblue")
fa("rocket", fill = "steelblue")
```


</font>
</center>
```` 



Which renders as:

<center>
<font size="15"> 
```{r fa2, message=FALSE, echo = FALSE, results='hold'}
library(fontawesome)
fa("r-project", fill = "steelblue")
fa("bolt-lightning", fill = "steelblue")
fa("discourse", fill = "steelblue")
fa("rocket", fill = "steelblue")
```


</font>
</center>


Note the added benefit of different font size and centering on the page.



# Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
devtools::session_info()

```


